---
output: rmarkdown::github_document
---

# Session 3 - Quantitative Tutorials

## Week 4 - ANOVA II

Dr. Erola Fenollosa (she/her)

\*\*/1/2023


### 1. Introduction

The aim of this tutorial is to perform and understand your first two-way ANOVA. A two-way ANOVA is nothing more than an ANOVA with two factors, this is two categorical variables. It is a very useful tool when we want to evaluate the influence of two categorical variables and their interaction on a numerical variable. 

For example, we performed an experiment of monitoring nutrient addition responses on plants and evaluate their photosynthetic rate at different times. We expect that the treatment (control / adding nutrients) factor affects the photosynthetic rate and that this effect gets bigger with time, so when we measure plants after two weeks we see higher differences between treatments than after one week. In this example we have two factors: Treatment (with two levels: control and adding nutrients) and Time (with three levels: one week, two weeks, three weeks). Those factors are crossed because every category (level) of one factor co-occurs in the design with every category of the other factor, in other words, all times have information on all treatments and all treatments information on all times. 

When we have a two-way crossed ANOVA we have a new element: the interaction. In ANOVA, the interaction is defined as when the difference in the means of the response between the levels of one factor is not the same across all levels of another factor. For example, if the effect of the treatment increases with time it means the interaction is significant, so the effect of a factor is influenced by the other.

In this script you'll find a compendium of exercises to code and explain the different required steps to code a two-way anova. Fill all exercises including comments when required, a "TO DO" will help you identify the parts to fill.



### 2. Import data and load packages
Open Top chambers

Temperature
Altered precipitation
One species one time

Include pictures


```{r}
# Load packages
library(readxl)

```


```{r}
# Data import (ensure the csv file is at the same folder as this Rmd)


```


### 3. Research question

TODO 3.1: The research question we want to solve is: ... ?


TODO 3.2: The hypothesis are:



To solve this hypothesis we will perform a two way ANOVA with the fixed factors: Temperature and Rainfall on the continuous variable of seedling height (cm).


### 4. Prepare the data
After describing the research question and hypotheses let's prepare the data for the analysis.

```{r}
# Ensure your factors are actually factors for R
# <- as.factor()
#<- as.factor()
```


```{r}
# Check if we have missing values
```

```{r}
# Get a general overview of the data
#summary()
```

```{r}
# Visualize the results to see if our hypotheses make sense

```


### 5. ANOVA Assumptions
Let's check if our data meet ANOVA requirements: data normality (our continuous variable data is normally distributed) and homocedasticity (groups have equal variances).


```{r}
# TODO 5.1: Check normal distribution
  
```


```{r}
# TODO 5.2: Check homocedasticity 
  
```

TODO 5.3. What results did you obtain?  What do they mean?


### 6. Two-way fixed crossed ANOVA
We are going to perform now the two-way ANOVA:
```{r}

```

TODO 6.1. What is this line of code above doing? Explain it with your own words

TODO 6.2. Is there any p-value under 0.05? What does it mean?


### 7. Post hoc

```{r}
# TODO 7.1: FOR Temperature
```

```{r}
# TODO 7.2: FOR Rainfall
```

```{r}
# TODO 7.3: Interaction
```

TODO 7.4. Is there any p-value under 0.05? What does it mean?


### 8. Further questions
- How would you compare if different species respond differently?
- And how would you compare the effect of time?

### 9. Conclusions

TODO 9.1 Write down the conclusions of this work. Expose if the ANOVA requirements were meet, the ANOVA results, their meaning, if we should accept the hypotheses or not, and specially what does this result mean for the climate change research.






